---
title: "Modeltime: Demo"
author: "Sergio Cubero"
date: "1/4/2022"
output: html_document
---

Ejemplo tomado: 

```{r ,echo=FALSE}
vembedr::embed_youtube("6RjYIOCnRMk")
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r paquetes}
# Machine Learning
library(tidymodels)
library(modeltime)
library(modeltime.ensemble)
library(modeltime.resample)

# Time Series
library(timetk)

# Core
library(tidyverse)
```

## DATA

```{r}
walmart_sales_weekly

walmart_sales_weekly%>%
  group_by(id)%>%
  plot_time_series(Date,Weekly_Sales,.facet_ncol = 3, .interactive = F)
```

## DATA PREPARATION

```{r}
FORECAST_HORIZON <- 52
```

### Full = Training + Forecast Dataset

```{r}
full_data_tbl <- walmart_sales_weekly%>%
  select(id,Date,Weekly_Sales)%>%
  
  ## Apply Group-Wise Time Series Manipulations
  group_by(id)%>%
  future_frame(
    .date_var = Date,
    .length_out = FORECAST_HORIZON,
    .bind_data = T
  )%>%
  ungroup()%>%
  
  # Consolidate IDs
  mutate(id=fct_drop(id))

```

### Training Data

```{r}
data_prepared_tbl <- full_data_tbl%>%
  filter(!is.na(Weekly_Sales))
  
data_prepared_tbl%>%
  group_by(id)%>%
  tk_summary_diagnostics()
```

### Future Data Forecast
```{r}
future_tbl <- full_data_tbl%>%
  filter(is.na(Weekly_Sales))
```

## SPLITTING

```{r}
splits <- data_prepared_tbl%>%
  arrange(id,Date)%>%
  time_series_split(
    data_var=Date,
    assess = FORECAST_HORIZON,
    cumulative = T
  )

splits

# Testing Count
84/7

# Training Count
917/7
```

## PREPROCESOR

```{r}
recipe_spec_1 <- recipe(Weekly_Sales~., training(splits))%>%
  step_timeseries_signature(Date)%>%
  ## Elimina las columnas o atributos que no aportan
  step_rm(matches("(.iso$)|(.xts)|(day)|(hour)|(minute)|(second)|(am.pm)")) %>%
  step_normalize(Date_index.num,Date_year)%>%
  step_mutate(Date_week = factor(Date_week,ordered = T))%>%
  step_dummy(all_nominal(),one_hot = T)

recipe_spec_1 %>% prep() %>% juice() %>% glimpse()

recipe_spec_2 <- recipe_spec_1%>%
  update_role(Date,new_role = "ID")

recipe_spec_2 %>% prep() %>% juice() %>% glimpse()

recipe_spec_1 %>% prep() %>% summary()
recipe_spec_2 %>% prep() %>% summary()
```

## MODELS

### prophet 

```{r}
wflw_fit_prophet <- workflow()%>%
  add_model(
    prophet_reg() %>% set_engine("prophet")
  )%>%
  add_recipe(recipe_spec_1)%>%
  fit(training(splits))
```

### XGBOOST

```{r}
wflw_fit_xgboost <- workflow()%>%
  add_model(
    boost_tree() %>% set_engine("xgboost")
  )%>%
  add_recipe(recipe_spec_2)%>%
  fit(training(splits))
```

### Random Forest

```{r}
wflw_fit_rf <- workflow()%>%
  add_model(
    rand_forest() %>% set_engine("ranger")
  )%>%
  add_recipe(recipe_spec_2)%>%
  fit(training(splits))
```

### SVM

```{r}
wflw_fit_svm <- workflow()%>%
  add_model(
    svm_rbf() %>% set_engine("kernlab")
  )%>%
  add_recipe(recipe_spec_2)%>%
  fit(training(splits))
```

### prophet_boost

```{r}
wflw_fit_prophet_boost <- workflow()%>%
  add_model(
    prophet_boost(
      seasonality_yearly = F,
      seasonality_weekly = F,
      seasonality_daily =  F,
    ) %>% 
      set_engine("prophet_xgboost")
  )%>%
  add_recipe(recipe_spec_1)%>%
  fit(training(splits))
```

## MODELTIME WORKFLOW

### modeltime table

```{r}
submodels_tbl <- modeltime_table(
  wflw_fit_prophet,
  wflw_fit_prophet_boost,
  wflw_fit_xgboost,
  wflw_fit_rf,
  wflw_fit_svm
)

submodels_tbl
```

### calibrate Testing Data

```{r}
submodels_calibrated_tbl <- submodels_tbl %>%
  modeltime_calibrate(testing(splits))

submodels_calibrated_tbl
```

### Measure Test Accuracy

```{r}
submodels_calibrated_tbl%>% modeltime_accuracy()
```

### Visualize test forecast

```{r}
submodels_calibrated_tbl %>%
  modeltime_forecast(
    new_data = testing(splits),
    actual_data = data_prepared_tbl,
    keep_data = T
  )%>%
  group_by(id)%>%
  plot_modeltime_forecast(
    .facet_ncol=2
  )
```

### Refit on full training dataset

```{r}
submodels_refit_tbl <- submodels_calibrated_tbl %>%
  modeltime_refit(data_prepared_tbl)
```

### Visualize Submodel Forecast

```{r}
submodels_refit_tbl%>%
  modeltime_forecast(
    new_data =  future_tbl,
    actual_data = data_prepared_tbl,
    keep_data = T
  )%>%
  group_by(id) %>%
  plot_modeltime_forecast(
    .facet_ncol=2
  )
```

## ENSEMBLE

## Make ensemble

```{r}
ensemble_fit_mean <- submodels_tbl %>%
  filter(!.model_id %in% c(1))%>%
  ensemble_average(type="mean")

# Modeltime table
ensemble_tbl <- modeltime_table(
  ensemble_fit_mean
)

ensemble_tbl
```

### Ensemble test Accuracy

```{r}
ensemble_tbl%>%
  combine_modeltime_tables(submodels_tbl)%>%
  modeltime_accuracy(testing(splits))
```

### Ensemble Test Forecast

```{r}
ensemble_tbl%>%
  modeltime_forecast(
    new_data =  testing(splits),
    actual_data = data_prepared_tbl,
    keep_data = T
  )%>%
  group_by(id) %>%
  plot_modeltime_forecast(
    .facet_ncol=2
  )
```

### Refit Ensemble

```{r}
ensemble_refit_tbl <- ensemble_tbl%>%
  modeltime_refit(data_prepared_tbl)
```

### Visualize Ensemble Forecast

```{r}
ensemble_refit_tbl%>%
  modeltime_forecast(
    new_data =  future_tbl,
    actual_data = data_prepared_tbl,
    keep_data = T
  )%>%
  group_by(id) %>%
  plot_modeltime_forecast(
    .facet_ncol=2
  )

```
