---
title: "No lineales"
author: "Sergio Cubero"
date: "9/17/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(forecast)
library(TSA)
library(NTS)
library(MSwM)
library(tsDyn)
library(fNonlinear)
library(dlm)
library(astsa)
library(tidyverse)
library(readxl)
```


```{r datos, include=FALSE}

SeriesDatos <- read_excel("~/Google Drive/Mi unidad/1.Maestria II Ciclo 2021/Curso de Analisis De Casos/Caso II/Datos/Base Datos.xlsx")%>%
  janitor::clean_names()%>%
  mutate(ActivoNeto=paste0(activo_neto,"-01"))%>%
  rename('ActNetCRC'=crc,
         'ActNetUSD'=usd)

actnetcrc<- ts(SeriesDatos[,2],start =c(2001,2),end=c(2021,7), frequency = 12)
actnetusd<- ts(SeriesDatos[,3],start =c(2001,2),end=c(2021,7), frequency = 12)
actnet <- cbind(actnetcrc,actnetusd) 

## Peridos de Tiempo
inicio_train<- c(2011,1)
fin_train<- c(2020,12)

inicio_test <- c(2021,1)

sactnetcrc<- window(actnetcrc,start=inicio_train)
sactnetcrc_train<- window(actnetcrc,start=inicio_train, end=fin_train)
sactnetcrc_test<- window(actnetcrc,start=inicio_test)

sactnetusd<- window(actnetusd,start=inicio_train)
sactnetusd_train<- window(actnetusd,start=inicio_train, end=fin_train)
sactnetusd_test<- window(actnetusd,start=inicio_test)
```



# Pruebas de No Linealidad Colones

## Teraesvirta Neural Network Test
```{r}
fNonlinear::tnnTest(log(sactnetcrc_train), lag = 1, title = NULL, description = NULL)
```

## Kennan tests for nonlineary

La hipótesis nula de que la serie de tiempo sigue algún proceso de AR.

```{r}
Keenan.test(log(sactnetcrc_train))
Keenan.test(log(sactnetcrc_train), order=1)
Keenan.test(log(sactnetcrc_train), order=2)
Keenan.test(log(sactnetcrc_train), order=3)
```


## BDS

```{r}
bdsTest(log(sactnetcrc_train))
```


## Thresold Non Linearity

```{r}
lag1.plot(log(sactnetcrc_train), max.lag=12)
thr.test(log(sactnetcrc_train))
thr.test(log(sactnetcrc_train),p=2,d=1)
thr.test(log(sactnetcrc_train),p=2,d=2)
thr.test(log(sactnetcrc_train),p=3,d=1)
thr.test(log(sactnetcrc_train),p=3,d=2)
thr.test(log(sactnetcrc_train),p=3,d=3)
```


# Modelos No Lineales

## TAR

```{r ,eval=F}

set.seed(777)
# m orden
pm <- 1:4

mod.list.tar<-list()



for(j in pm){
  for(i in pm){
    for(l in pm){
      mod<-TSA::tar(log(sactnetcrc_train),p1=j,p2=i,d=l)  
      
      mod.list.tar[[paste(j,l,i,sep="-")]]<-mod$AIC
      print(paste(j,l,i,sep="-"))    
    }
  }
}

AICTar<-bind_rows(mod.list.tar,.id = "Ordene-delay")%>%
  arrange(`1`)

knitr::kable(head(AICTar,20))

```

```{r}
mod.tar1<-TSA::tar(log(sactnetcrc_train),p1=2,p2=2,d=1)  
mod.tar2<-TSA::tar(log(sactnetcrc_train),p1=3,p2=3,d=1)  
mod.tar3<-TSA::tar(log(sactnetcrc_train),p1=3,p2=4,d=3)  

# mod.tar1$thd
# mod.tar2$thd
mod.tar3$thd

# mod.tar1$qr1$coefficients
# mod.tar2$qr1$coefficients
mod.tar3$qr1$coefficients

# mod.tar1$qr2$coefficients
# mod.tar2$qr2$coefficients
mod.tar3$qr2$coefficients

mod.tar1$AIC
mod.tar2$AIC
mod.tar3$AIC


# checkresiduals(ts(mod.tar1$residuals,start=c(2011,1),frequency = 12))
# checkresiduals(ts(mod.tar2$residuals,start=c(2011,1),frequency = 12))
checkresiduals(ts(mod.tar3$residuals,start=c(2011,1),frequency = 12))
```


```{r}
# prontar1<- ts(as.vector(predict(mod.tar1,n.ahead=7,n.sim=1000)$fit),start=c(2021,1),frequency = 12)
# prontar2<- ts(as.vector(predict(mod.tar2,n.ahead=7,n.sim=1000)$fit),start=c(2021,1),frequency = 12)
prontar3<- ts(as.vector(predict(mod.tar3,n.ahead=7,n.sim=1000)$fit),start=c(2021,1),frequency = 12)

# fit1<-ts(exp(log(sactnetcrc_train)-mod.tar1$residuals),start =c(2011,1),frequency = 12)
# fit2<-ts(exp(log(sactnetcrc_train)-mod.tar2$residuals),start =c(2011,1),frequency = 12)
fit3<-ts(exp(log(sactnetcrc_train)-mod.tar3$residuals),start =c(2011,1),frequency = 12)

autoplot(sactnetcrc_train)+
  # autolayer(fit1)+
  # autolayer(fit2)+
  autolayer(fit3)+
  theme_bw()

Metrics::rmse(sactnetcrc_test, exp(prontar1))
Metrics::rmse(sactnetcrc_test, exp(prontar2))
Metrics::rmse(sactnetcrc_test, exp(prontar3))

autoplot(sactnetcrc_test)+
  # autolayer(exp(prontar1))+
  # autolayer(exp(prontar2))+
  autolayer(exp(prontar3))+
  theme_bw()
```

## SETAR

Thus the threshold delay, the number of lags in each regime and the threshold value are computed.

```{r, eval=F}

n2<- selectSETAR(log(sactnetcrc_train),m=3,thDelay=seq(1,2,by=1),nthresh=2,criterion = "SSR",type="level")
n11<-selectSETAR(log(sactnetcrc_train),m=3,thDelay=seq(0,2,by=1),nthresh=1,d=1,type="level")

n2$bests2th
n11$bests


```

```{r}
set.seed(777)
modeloas1<-setar(log(sactnetcrc_train),m=3,mL=3,mH=1,nthresh=1,thDelay = 2,type="level")
# summary(modeloas1) #residuals variance = 0.005525,  AIC = -632, MAPE = 0.4352%
#plot(modeloas1)

modeloas2<-setar(log(sactnetcrc_train),m=3,mL=3,mH=2,nthresh=1,thDelay = 2,type="level")
summary(modeloas2) # residuals variance = 0.005857,  AIC = -635, MAPE = 0.4584%
#plot(modeloas2)

modeloas3<-setar(log(sactnetcrc_train),m=3,d=1,thDelay = 1,nthresh=2,th=c(13.144564,13.494865),type="level")
# summary(modeloas3) # residuals variance = 0.006319,  AIC = -621, MAPE = 0.4621%
#plot(modeloas3)

modeloas4<-setar(log(sactnetcrc_train),m=2,mL=2,mH=2,nthresh=1,thDelay = 0,type="level",th=13.54936)
# summary(modeloas4) # residuals variance = 0.006319,  AIC = -621, MAPE = 0.4621%
#plot(modeloas4)


AIC(modeloas1)
AIC(modeloas2)
AIC(modeloas3)
AIC(modeloas4)
```

```{r}
# pronsetar1<- predict(modeloas1, n.ahead = 7)
pronsetar2<- predict(modeloas2, n.ahead = 7)
# pronsetar3<- predict(modeloas3, n.ahead = 7)
# pronsetar4<- predict(modeloas4, n.ahead = 7)

# fit1<-ts(exp(modeloas1$fitted.values),start =c(2011,1),frequency = 12)
fit2<-ts(exp(modeloas2$fitted.values),start =c(2011,1),frequency = 12)
# fit3<-ts(exp(modeloas3$fitted.values),start =c(2011,1),frequency = 12)
# fit4<-ts(exp(modeloas4$fitted.values),start =c(2011,1),frequency = 12)

autoplot(sactnetcrc_train)+
  # autolayer(fit1)+
  autolayer(fit2)+
  # autolayer(fit3)+
  # autolayer(fit4)+
  theme_bw()

Metrics::rmse(sactnetcrc_test, exp(pronsetar1))
Metrics::rmse(sactnetcrc_test, exp(pronsetar2))
Metrics::rmse(sactnetcrc_test, exp(pronsetar3))
Metrics::rmse(sactnetcrc_test, exp(pronsetar4))

autoplot(sactnetcrc_test)+
  # autolayer(exp(pronsetar1))+
  autolayer(exp(pronsetar2))+
  # autolayer(exp(pronsetar3))+
  # autolayer(exp(pronsetar4))+
  theme_bw()




```


## LSTAR

```{r}

modlstat1<-lstar(log(sactnetcrc_train), m=1, include = c("const", "trend","none", "both"))     
modlstat2<-lstar(log(sactnetcrc_train), m=2, include = c("const", "trend","none", "both"))  
modlstat3<-lstar(log(sactnetcrc_train), m=3, include = c("const", "trend","none", "both"))  

# summary(modlstat1)
# summary(modlstat2)
summary(modlstat3)

pronlsetar1<- predict(modlstat1, n.ahead = 7)
pronlsetar2<- predict(modlstat2, n.ahead = 7)
pronlsetar3<- predict(modlstat3, n.ahead = 7)

Metrics::rmse(sactnetcrc_test, exp(pronlsetar1))
Metrics::rmse(sactnetcrc_test, exp(pronlsetar2))
Metrics::rmse(sactnetcrc_test, exp(pronlsetar3))

# fit1<-ts(exp(modlstat1$fitted.values),start =c(2011,1),frequency = 12)
# fit2<-ts(exp(modlstat2$fitted.values),start =c(2011,1),frequency = 12)
fit3<-ts(exp(modlstat3$fitted.values),start =c(2011,1),frequency = 12)


autoplot(sactnetcrc_train)+
  # autolayer(fit1)+
  # autolayer(fit2)+
  autolayer(fit3)+
  theme_bw()


autoplot(sactnetcrc_test)+
  # autolayer(exp(pronlsetar1))+
  # autolayer(exp(pronlsetar2))+
  autolayer(exp(pronlsetar3))+
  theme_bw()

```

# Metricas Generales

```{r}
Metrics::rmse(sactnetcrc_test, exp(prontar3))
Metrics::rmse(sactnetcrc_test, exp(pronsetar2))
Metrics::rmse(sactnetcrc_test, exp(pronlsetar3))

autoplot(sactnetcrc_test)+
  autolayer( exp(prontar3))+
  autolayer(exp(pronsetar2))+
  autolayer(exp(pronlsetar3))+
  theme_bw()
```

# Pruebas de No Linealidad Dolares

## Teraesvirta Neural Network Test
```{r}
fNonlinear::tnnTest(log(sactnetusd_train), lag = 1, title = NULL, description = NULL)
```

## Kennan tests for nonlineary

La hipótesis nula de que la serie de tiempo sigue algún proceso de AR.

```{r}
Keenan.test(log(sactnetcrc_train))
Keenan.test(log(sactnetusd_train), order=1)
Keenan.test(log(sactnetusd_train), order=2)
Keenan.test(log(sactnetusd_train), order=3)
```


## BDS

```{r}
bdsTest(log(sactnetusd_train))
```


## Thresold Non Linearity

```{r}
lag1.plot(log(sactnetusd_train), max.lag=12)
thr.test(log(sactnetusd_train))
thr.test(log(sactnetusd_train),p=2,d=1)
thr.test(log(sactnetusd_train),p=2,d=2)
thr.test(log(sactnetusd_train),p=3,d=1)
thr.test(log(sactnetusd_train),p=3,d=2)
thr.test(log(sactnetusd_train),p=3,d=3)
```


# Modelos No Lineales

## TAR

```{r ,eval=F}

set.seed(777)
# m orden
pm <- 1:4

mod.list.tar<-list()



for(j in pm){
  for(i in pm){
    for(l in pm){
      mod<-TSA::tar(log(sactnetusd_train),p1=j,p2=i,d=l)  
      
      mod.list.tar[[paste(j,l,i,sep="-")]]<-mod$AIC
      print(paste(j,l,i,sep="-"))    
    }
  }
}

AICTar<-bind_rows(mod.list.tar,.id = "Ordene-delay")%>%
  arrange(`1`)

knitr::kable(head(AICTar,20))

```

```{r}
mod.tar1<-TSA::tar(log(sactnetusd_train),p1=2,p2=2,d=1)  
mod.tar2<-TSA::tar(log(sactnetusd_train),p1=3,p2=3,d=1)  
mod.tar3<-TSA::tar(log(sactnetusd_train),p1=3,p2=4,d=3)  

# mod.tar1$thd
# mod.tar2$thd
mod.tar3$thd

# mod.tar1$qr1$coefficients
# mod.tar2$qr1$coefficients
mod.tar3$qr1$coefficients

# mod.tar1$qr2$coefficients
# mod.tar2$qr2$coefficients
mod.tar3$qr2$coefficients

mod.tar1$AIC
mod.tar2$AIC
mod.tar3$AIC


# checkresiduals(ts(mod.tar1$residuals,start=c(2011,1),frequency = 12))
# checkresiduals(ts(mod.tar2$residuals,start=c(2011,1),frequency = 12))
checkresiduals(ts(mod.tar3$residuals,start=c(2011,1),frequency = 12))
```


```{r}
# prontar1<- ts(as.vector(predict(mod.tar1,n.ahead=7,n.sim=1000)$fit),start=c(2021,1),frequency = 12)
# prontar2<- ts(as.vector(predict(mod.tar2,n.ahead=7,n.sim=1000)$fit),start=c(2021,1),frequency = 12)
prontar3<- ts(as.vector(predict(mod.tar3,n.ahead=7,n.sim=1000)$fit),start=c(2021,1),frequency = 12)

# fit1<-ts(exp(log(sactnetusd_train)-mod.tar1$residuals),start =c(2011,1),frequency = 12)
# fit2<-ts(exp(log(sactnetusd_train)-mod.tar2$residuals),start =c(2011,1),frequency = 12)
fit3<-ts(exp(log(sactnetusd_train)-mod.tar3$residuals),start =c(2011,1),frequency = 12)

autoplot(sactnetusd_train)+
  # autolayer(fit1)+
  # autolayer(fit2)+
  autolayer(fit3)+
  theme_bw()

Metrics::rmse(sactnetusd_test, exp(prontar1))
Metrics::rmse(sactnetusd_test, exp(prontar2))
Metrics::rmse(sactnetusd_test, exp(prontar3))

autoplot(sactnetusd_test)+
  # autolayer(exp(prontar1))+
  # autolayer(exp(prontar2))+
  autolayer(exp(prontar3))+
  theme_bw()
```

## SETAR

Thus the threshold delay, the number of lags in each regime and the threshold value are computed.

```{r, eval=F}

n2<- selectSETAR(log(sactnetusd_train),m=3,thDelay=seq(1,2,by=1),nthresh=2,criterion = "SSR",type="level")
n11<-selectSETAR(log(sactnetusd_train),m=3,thDelay=seq(0,2,by=1),nthresh=1,d=1,type="level")

n2$bests2th
n11$bests


```

```{r}
set.seed(777)
modeloas1<-setar(log(sactnetusd_train),m=3,mL=3,mH=1,nthresh=1,thDelay = 2,type="level")
# summary(modeloas1) #residuals variance = 0.005525,  AIC = -632, MAPE = 0.4352%
#plot(modeloas1)

modeloas2<-setar(log(sactnetusd_train),m=3,mL=3,mH=2,nthresh=1,thDelay = 2,type="level")
summary(modeloas2) # residuals variance = 0.005857,  AIC = -635, MAPE = 0.4584%
#plot(modeloas2)

modeloas3<-setar(log(sactnetusd_train),m=3,d=1,thDelay = 1,nthresh=2,th=c(13.144564,13.494865),type="level")
# summary(modeloas3) # residuals variance = 0.006319,  AIC = -621, MAPE = 0.4621%
#plot(modeloas3)

modeloas4<-setar(log(sactnetusd_train),m=2,mL=2,mH=2,nthresh=1,thDelay = 0,type="level",th=13.54936)
# summary(modeloas4) # residuals variance = 0.006319,  AIC = -621, MAPE = 0.4621%
#plot(modeloas4)


AIC(modeloas1)
AIC(modeloas2)
AIC(modeloas3)
AIC(modeloas4)
```

```{r}
# pronsetar1<- predict(modeloas1, n.ahead = 7)
pronsetar2<- predict(modeloas2, n.ahead = 7)
# pronsetar3<- predict(modeloas3, n.ahead = 7)
# pronsetar4<- predict(modeloas4, n.ahead = 7)

# fit1<-ts(exp(modeloas1$fitted.values),start =c(2011,1),frequency = 12)
fit2<-ts(exp(modeloas2$fitted.values),start =c(2011,1),frequency = 12)
# fit3<-ts(exp(modeloas3$fitted.values),start =c(2011,1),frequency = 12)
# fit4<-ts(exp(modeloas4$fitted.values),start =c(2011,1),frequency = 12)

autoplot(sactnetusd_train)+
  # autolayer(fit1)+
  autolayer(fit2)+
  # autolayer(fit3)+
  # autolayer(fit4)+
  theme_bw()

Metrics::rmse(sactnetusd_test, exp(pronsetar1))
Metrics::rmse(sactnetusd_test, exp(pronsetar2))
Metrics::rmse(sactnetusd_test, exp(pronsetar3))
Metrics::rmse(sactnetusd_test, exp(pronsetar4))

autoplot(sactnetusd_test)+
  # autolayer(exp(pronsetar1))+
  autolayer(exp(pronsetar2))+
  # autolayer(exp(pronsetar3))+
  # autolayer(exp(pronsetar4))+
  theme_bw()




```


## LSTAR

```{r}

modlstat1<-lstar(log(sactnetusd_train), m=1, include = c("const", "trend","none", "both"))     
modlstat2<-lstar(log(sactnetusd_train), m=2, include = c("const", "trend","none", "both"))  
modlstat3<-lstar(log(sactnetusd_train), m=3, include = c("const", "trend","none", "both"))  

# summary(modlstat1)
# summary(modlstat2)
summary(modlstat3)

pronlsetar1<- predict(modlstat1, n.ahead = 7)
pronlsetar2<- predict(modlstat2, n.ahead = 7)
pronlsetar3<- predict(modlstat3, n.ahead = 7)

Metrics::rmse(sactnetusd_test, exp(pronlsetar1))
Metrics::rmse(sactnetusd_test, exp(pronlsetar2))
Metrics::rmse(sactnetusd_test, exp(pronlsetar3))

# fit1<-ts(exp(modlstat1$fitted.values),start =c(2011,1),frequency = 12)
# fit2<-ts(exp(modlstat2$fitted.values),start =c(2011,1),frequency = 12)
fit3<-ts(exp(modlstat3$fitted.values),start =c(2011,1),frequency = 12)


autoplot(sactnetusd_train)+
  # autolayer(fit1)+
  # autolayer(fit2)+
  autolayer(fit3)+
  theme_bw()


autoplot(sactnetusd_test)+
  # autolayer(exp(pronlsetar1))+
  # autolayer(exp(pronlsetar2))+
  autolayer(exp(pronlsetar3))+
  theme_bw()

```

# Metricas Generales

```{r}
Metrics::rmse(sactnetusd_test, exp(prontar3))
Metrics::rmse(sactnetusd_test, exp(pronsetar2))
Metrics::rmse(sactnetusd_test, exp(pronlsetar3))

autoplot(sactnetusd_test)+
  autolayer( exp(prontar3))+
  autolayer(exp(pronsetar2))+
  autolayer(exp(pronlsetar3))+
  theme_bw()
```

