---
title: "No lineales"
author: "Sergio Cubero"
date: "9/17/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(forecast)
library(TSA)
library(NTS)
library(MSwM)
library(tsDyn)
library(fNonlinear)
library(dlm)
library(astsa)
library(tidyverse)
library(readxl)
```

## Separación de datos
```{r datos, include=FALSE}

SeriesDatos <- read_excel("~/Google Drive/Mi unidad/1.Maestria II Ciclo 2021/Curso de Analisis De Casos/Caso II/Datos/Base Datos.xlsx")%>%
  janitor::clean_names()%>%
  mutate(ActivoNeto=paste0(activo_neto,"-01"))%>%
  rename('ActNetCRC'=crc,
         'ActNetUSD'=usd)

actnetcrc<- ts(SeriesDatos[,2],start =c(2001,2),end=c(2021,7), frequency = 12)
actnetusd<- ts(SeriesDatos[,3],start =c(2001,2),end=c(2021,7), frequency = 12)
actnet <- cbind(actnetcrc,actnetusd) 

## Peridos de Tiempo
inicio_train<- c(2011,1)
fin_train<- c(2020,12)

inicio_test <- c(2021,1)

sactnetcrc<- window(actnetcrc,start=inicio_train)
sactnetcrc_train<- window(actnetcrc,start=inicio_train, end=fin_train)
sactnetcrc_test<- window(actnetcrc,start=inicio_test)

sactnetusd<- window(actnetusd,start=inicio_train)
sactnetusd_train<- window(actnetusd,start=inicio_train, end=fin_train)
sactnetusd_test<- window(actnetusd,start=inicio_test)
```

# Pruebas de No Linealidad Colones

## Colones

### General

```{r}
nonlinearTseries::nonlinearityTest(actnetcrc, verbose = TRUE)
```

### Teraesvirta Neural Network Test
```{r}
fNonlinear::tnnTest(actnetcrc, lag = 1, title = NULL, description = NULL)
```

### Kennan tests for nonlineary

La hipótesis nula de que la serie de tiempo sigue algún proceso de AR.

```{r}
Keenan.test(actnetcrc)
Keenan.test(actnetcrc, order=1)
Keenan.test(actnetcrc, order=2)
Keenan.test(actnetcrc, order=3)
```


### BDS

```{r}
bdsTest(actnetcrc)
```


### Thresold Non Linearity

```{r}
lag1.plot(actnetcrc, max.lag=12)
thr.test(actnetcrc)
thr.test(actnetcrc,p=2,d=1)
thr.test(actnetcrc,p=2,d=2)
thr.test(actnetcrc,p=3,d=1)
thr.test(actnetcrc,p=3,d=2)
thr.test(actnetcrc,p=3,d=3)
```

## Dolares

### General

```{r}
nonlinearTseries::nonlinearityTest(actnetusd, verbose = TRUE)
```


### Teraesvirta Neural Network Test
```{r}
fNonlinear::tnnTest(actnetusd, lag = 1, title = NULL, description = NULL)
```

### Kennan tests for nonlineary

La hipótesis nula de que la serie de tiempo sigue algún proceso de AR.

```{r}
Keenan.test(actnetusd)
Keenan.test(actnetusd, order=1)
Keenan.test(actnetusd, order=2)
Keenan.test(actnetusd, order=3)
```


### BDS

```{r}
bdsTest(actnetusd)
```


### Thresold Non Linearity

```{r}
lag1.plot(actnetusd, max.lag=12)
thr.test(actnetusd)
thr.test(actnetusd,p=2,d=1)
thr.test(actnetusd,p=2,d=2)
thr.test(actnetusd,p=3,d=1)
thr.test(actnetusd,p=3,d=2)
thr.test(actnetusd,p=3,d=3)
```



# Modelos No Lineales

## Colones

### TAR

```{r, include=F}
rm(list=ls()[!ls() %in% c('actnetcrc','actnetusd','actnet',"fin_train","inicio_test","inicio_train","sactnetcrc","sactnetcrc_test","sactnetcrc_train","sactnetusd","sactnetusd_test","sactnetusd_train")])
```

```{r ,eval=F}

# m orden
pm <- 1:3

mod.list.tar<-list()
AIC.best.list<-list()

AICM = NULL
model.best <- list(d=0, p1=0, p2=0)
AIC.best = 2888

for(l in pm){
  for(j in pm){
    for(i in pm){
      set.seed(777)
      model.tar.s = tar(sactnetcrc_train,p1=j,p2=i,d=l)
      mod.list.tar[[paste(j,i,l,sep="-")]]<-model.tar.s$AIC
      #print(paste(j,i,l,sep="-"))    
      
      if (model.tar.s$AIC < AIC.best) {
		    AIC.best = model.tar.s$AIC
		    AIC.best.list[[paste(j,i,l,sep="-")]]<-AIC.best
		    #print(AIC.best)
		    model.best$d = l
		    model.best$p1 = model.tar.s$p1
		    model.best$p2 = model.tar.s$p2 
		    print(paste(model.tar.s$p1,model.tar.s$p2,l,sep="-")) }
    }
  }
}

# AICTar<-bind_rows(mod.list.tar,.id = "Ordene-delay")%>%
#   arrange(`1`)
# 
# knitr::kable(head(AICTar,20))

AICTarBest<-bind_rows(AIC.best.list,.id = "Ordene-delay")%>%
  arrange(`1`)

knitr::kable(head(AICTarBest,20))

```

```{r}
mod.tar1<-TSA::tar(sactnetcrc_train,p1=2,p2=3,d=1)  
mod.tar2<-TSA::tar(sactnetcrc_train,p1=3,p2=1,d=1)  
mod.tar3<-TSA::tar(sactnetcrc_train,p1=3,p2=2,d=1)  

mod.tar1$thd
mod.tar2$thd
mod.tar3$thd

mod.tar1$qr1$coefficients
mod.tar2$qr1$coefficients
mod.tar3$qr1$coefficients

mod.tar1$qr2$coefficients
mod.tar2$qr2$coefficients
mod.tar3$qr2$coefficients

cbind(
Modelo=c("p1=2,p2=3,d=1",
         "p1=3,p2=1,d=1",
         "p1=3,p2=2,d=1"),
AIC=c(mod.tar1$AIC,
mod.tar2$AIC,
mod.tar3$AIC))%>%
  knitr::kable()


#tsdiag(mod.tar1)
tsdiag(mod.tar2)
#tsdiag(mod.tar3)


checkresiduals(ts(mod.tar1$residuals,start=c(2011,1),frequency = 12))
checkresiduals(ts(mod.tar2$residuals,start=c(2011,1),frequency = 12))
checkresiduals(ts(mod.tar3$residuals,start=c(2011,1),frequency = 12))
```


```{r}
prontar1<- ts(as.vector(predict(mod.tar1,n.ahead=7,n.sim=1000)$fit),start=c(2021,1),frequency = 12)
prontar2<- ts(as.vector(predict(mod.tar2,n.ahead=7,n.sim=1000)$fit),start=c(2021,1),frequency = 12)
prontar3<- ts(as.vector(predict(mod.tar3,n.ahead=7,n.sim=1000)$fit),start=c(2021,1),frequency = 12)

fit1<-ts(as.vector(mod.tar1$y)-as.vector(mod.tar1$residuals),start =c(2011,1),frequency = 12)
fit2<-ts(sactnetcrc_train-mod.tar2$residuals,start =c(2011,1),frequency = 12)
fit3<-ts(sactnetcrc_train-mod.tar3$residuals,start =c(2011,1),frequency = 12)

autoplot(sactnetcrc_train)+
  autolayer(fit1)+
  autolayer(fit2)+
  autolayer(fit3)+
  theme_bw()

Metrics::rmse(sactnetcrc_test, prontar1)
Metrics::rmse(sactnetcrc_test, prontar2)
Metrics::rmse(sactnetcrc_test, prontar3)

autoplot(sactnetcrc_test)+
  autolayer(prontar1)+
  autolayer(prontar2)+
  autolayer(prontar3)+
  theme_bw()+
  scale_y_continuous(limits = c(500000,1400000))
```

### SETAR

Thus the threshold delay, the number of lags in each regime and the threshold value are computed.

```{r, eval=T}

Setar1 <-
  selectSETAR(
    sactnetcrc_train, 
    include = c("const", "trend","none", "both"),
    m = 3,
    thDelay = seq(1, 2, by = 1),
    nthresh = 2,
    criterion = "AIC",
    type = "level",
    plot = T,
    trace = T
  )

Setar2 <-
  selectSETAR(
    sactnetcrc_train,
    m = 3,
    d=2,
    thDelay = seq(1, 2, by = 1),
    nthresh = 2,
    criterion = "AIC",
    type = "level",
    plot = T,
    trace = T
  )

Setar3 <-
  selectSETAR(
    sactnetcrc_train,
    m = 3,
    thDelay = seq(0, 2, by = 1),
    nthresh = 1,
    d = 1,
    criterion = "AIC",
    type = "level",
    plot = T,
    trace = T
  )

Setar4 <-
  selectSETAR(
    sactnetcrc_train,
    m = 3,
    thDelay = seq(0, 2, by = 1),
    nthresh = 1,
    d = 2,
    criterion = "AIC",
    type = "level",
    plot = T,
    trace = T
  )


Setar1$allTh%>%
  as.data.frame()%>%
  arrange(AIC,thDelay,mL,mH)%>%
  head(5)


Setar2$allTh%>%
  as.data.frame()%>%
  arrange(AIC,thDelay,mL,mH)%>%
  head(5)

Setar3$allTh%>%
  as.data.frame()%>%
  arrange(AIC,thDelay,mL,mH)%>%
  head(5)

Setar4$allTh%>%
  as.data.frame()%>%
  arrange(AIC,thDelay,mL,mH)%>%
  head(5)
```

```{r Modelos Seleccionados}

modeloas1 <-
  setar(
    sactnetcrc_train,
    m = 3,
    mL = 3,
    mH = 1,
    d=1,
    nthresh = 1,
    thDelay = 2,
    type = "level"
  )
## Raiz Unitaria
summary(modeloas1) #residuals variance = 0.005525,  AIC = -632, MAPE = 0.4352%
# plot(modeloas1)
checkresiduals(ts(modeloas1$residuals,start=c(2011,1),frequency = 12))

modeloas2 <-
  setar(
    sactnetcrc_train,
    m = 3,
    mL = 2,
    mH = 3,
    d=2,
    nthresh = 1,
    thDelay = 2,
    type = "level"
  )
## Raiz Unitaria
summary(modeloas2) # residuals variance = 0.005857,  AIC = -635, MAPE = 0.4584%
# plot(modeloas2)
checkresiduals(ts(modeloas2$residuals,start=c(2011,1),frequency = 12))

modeloas3 <-
  setar(
    sactnetcrc_train,
    m = 3,
    mL = 3,
    mH = 2,
    d=1,
    nthresh = 1,
    thDelay = 0,
    type = "level"
  )
## Raiz Unitaria
summary(modeloas3) # residuals variance = 0.006319,  AIC = -621, MAPE = 0.4621%
# plot(modeloas3)
checkresiduals(ts(modeloas3$residuals,start=c(2011,1),frequency = 12))

modeloas4 <-
  setar(
    sactnetcrc_train,
    m = 3,
    mL = 1,
    mH = 2,
    d=2,
    nthresh = 1,
    thDelay = 0,
    type = "level"
  )
summary(modeloas4) # residuals variance = 0.006319,  AIC = -621, MAPE = 0.4621%
# plot(modeloas4)
checkresiduals(ts(modeloas4$residuals,start=c(2011,1),frequency = 12))


AIC(modeloas1)
AIC(modeloas2)
AIC(modeloas3)
AIC(modeloas4)
```

```{r}
pronsetar1<- predict(modeloas1, n.ahead = 7)
pronsetar2<- predict(modeloas2, n.ahead = 7)
pronsetar3<- predict(modeloas3, n.ahead = 7)
pronsetar4<- predict(modeloas4, n.ahead = 7)

fit1<-ts(modeloas1$fitted.values,start =c(2011,1),frequency = 12)
fit2<-ts(modeloas2$fitted.values,start =c(2011,1),frequency = 12)
fit3<-ts(modeloas3$fitted.values,start =c(2011,1),frequency = 12)
fit4<-ts(modeloas4$fitted.values,start =c(2011,1),frequency = 12)

autoplot(sactnetcrc_train)+
  autolayer(fit1)+
  autolayer(fit2)+
  autolayer(fit3)+
  autolayer(fit4)+
  theme_bw()

data.frame(
Modelo= c(
  "1) m = 3,mL = 3,mH = 1, d=1",
  "2) m = 3,mL = 2,mH = 3, d=2",
  "3) m = 3,mL = 3,mH = 2, d=1",
  "4) m = 3,mL = 1,mH = 2, d=2"
),
RMSE=c(
  Metrics::rmse(sactnetcrc_test, pronsetar1),
  Metrics::rmse(sactnetcrc_test, pronsetar2),
  Metrics::rmse(sactnetcrc_test, pronsetar3),
  Metrics::rmse(sactnetcrc_test, pronsetar4)))%>%
  arrange(RMSE)%>%
  knitr::kable()

autoplot(sactnetcrc_test)+
  autolayer(pronsetar1)+
  autolayer(pronsetar2)+
  autolayer(pronsetar3)+
  autolayer(pronsetar4)+
  theme_bw()+
  scale_y_continuous(limits = c(500000,1400000))
```

### Metricas Generales

```{r , eval=T}
Metrics::rmse(sactnetcrc_test,(prontar2))
Metrics::rmse(sactnetcrc_test, (pronsetar4))

autoplot(sactnetcrc_test)+
  autolayer(prontar2)+
  autolayer(pronsetar4)+
  theme_bw()+
  scale_y_continuous(limits = c(500000,1400000))
```

## Dolares

### TAR

```{r, include=F}
rm(list=ls()[!ls() %in% c('actnetcrc','actnetusd','actnet',"fin_train","inicio_test","inicio_train","sactnetcrc","sactnetcrc_test","sactnetcrc_train","sactnetusd","sactnetusd_test","sactnetusd_train","pronsetar4","modeloas4","prontar2","mod.tar2")])
```

```{r ,eval=F}
# m orden
pm <- 1:4

mod.list.tar<-list()
AIC.best.list<-list()

AICM = NULL
model.best <- list(d=0, p1=0, p2=0)
AIC.best = 10000

for(l in pm){
  for(j in pm){
    for(i in pm){
      set.seed(777)
      model.tar.s = tar(sactnetusd_train,p1=j,p2=i,d=l)
      mod.list.tar[[paste(j,i,l,sep="-")]]<-model.tar.s$AIC
      print(paste("Modelo:",j,i,l,sep="-"))    
      
      if (model.tar.s$AIC < AIC.best) {
		    AIC.best = model.tar.s$AIC
		    AIC.best.list[[paste(j,i,l,sep="-")]]<-AIC.best
		    #print("Modelo:",j,i,l,"AIC",AIC.best)
		    model.best$d = l
		    model.best$p1 = model.tar.s$p1
		    model.best$p2 = model.tar.s$p2 
		    print(paste(model.tar.s$p1,model.tar.s$p2,l,sep="-")) }
    }
  }
}

# AICTar<-bind_rows(mod.list.tar,.id = "Ordene-delay")%>%
#   arrange(`1`)
# 
# knitr::kable(head(AICTar,20))

AICTarBest<-bind_rows(AIC.best.list,.id = "Ordene-delay")%>%
  arrange(`1`)

knitr::kable(head(AICTarBest,20))
```

```{r}
mod.tar1.usd<-TSA::tar(sactnetusd_train,p1=3,p2=4,d=1)  
mod.tar2.usd<-TSA::tar(sactnetusd_train,p1=1,p2=2,d=1)  
mod.tar3.usd<-TSA::tar(sactnetusd_train,p1=1,p2=3,d=1)  

mod.tar1.usd$thd
mod.tar2.usd$thd
mod.tar3.usd$thd

mod.tar1.usd$qr1$coefficients
mod.tar2.usd$qr1$coefficients
mod.tar3.usd$qr1$coefficients

mod.tar1.usd$qr2$coefficients
mod.tar2.usd$qr2$coefficients
mod.tar3.usd$qr2$coefficients

data.frame(
Modelo=c("p1=3,p2=4,d=1",
         "p1=1,p2=2,d=1",
         "p1=1,p2=3,d=1"),
AIC=c(mod.tar1.usd$AIC,
mod.tar2.usd$AIC,
mod.tar3.usd$AIC))%>%
  arrange(AIC)%>%
  knitr::kable()


tsdiag(mod.tar1.usd)
tsdiag(mod.tar2.usd)
tsdiag(mod.tar3.usd)


checkresiduals(ts(mod.tar1.usd$residuals,start=c(2011,1),frequency = 12))
checkresiduals(ts(mod.tar2.usd$residuals,start=c(2011,1),frequency = 12))
checkresiduals(ts(mod.tar3.usd$residuals,start=c(2011,1),frequency = 12))
```


```{r}
prontar1.usd<- ts(as.vector(predict(mod.tar1.usd,n.ahead=7,n.sim=1000)$fit),start=c(2021,1),frequency = 12)
prontar2.usd<- ts(as.vector(predict(mod.tar2.usd,n.ahead=7,n.sim=1000)$fit),start=c(2021,1),frequency = 12)
prontar3.usd<- ts(as.vector(predict(mod.tar3.usd,n.ahead=7,n.sim=1000)$fit),start=c(2021,1),frequency = 12)

fit1.usd<-ts(as.vector(mod.tar1.usd$y)-as.vector(mod.tar1.usd$residuals),start =c(2011,1),frequency = 12)
fit2.usd<-ts(as.vector(mod.tar1.usd$y)-mod.tar2.usd$residuals,start =c(2011,1),frequency = 12)
fit3.usd<-ts(as.vector(mod.tar1.usd$y)-mod.tar3.usd$residuals,start =c(2011,1),frequency = 12)

autoplot(sactnetusd_train)+
  autolayer(fit1.usd)+
  autolayer(fit2.usd)+
  autolayer(fit3.usd)+
  theme_bw()

data.frame(
Modelo=c("p1=3,p2=4,d=1",
         "p1=1,p2=2,d=1",
         "p1=1,p2=3,d=1"),
RMSE=c(
  Metrics::rmse(sactnetusd_test, prontar1.usd),
  Metrics::rmse(sactnetusd_test, prontar2.usd),
  Metrics::rmse(sactnetusd_test, prontar3.usd)))%>%
  arrange(RMSE)%>%
  knitr::kable()



autoplot(sactnetusd_test)+
  autolayer(prontar1.usd)+
  autolayer(prontar2.usd)+
  autolayer(prontar3.usd)+
  theme_bw()
```

### SETAR

Thus the threshold delay, the number of lags in each regime and the threshold value are computed.

```{r, eval=t}
Setar1.usd <-
  selectSETAR(
    sactnetusd_train, 
    include = c("const", "trend","none", "both"),
    m = 4,
    thDelay = seq(0, 3, by = 1),
    nthresh = 3,
    criterion = "AIC",
    type = "level",
    plot = T,
    trace = T
  )

Setar2.usd <-
  selectSETAR(
    sactnetusd_train,
    m = 4,
    d=2,
    thDelay = seq(0, 3, by = 1),
    nthresh = 3,
    criterion = "AIC",
    type = "level",
    plot = T,
    trace = T
  )

Setar3.usd <-
  selectSETAR(
    sactnetusd_train,
    m = 4,
    thDelay = seq(0, 3, by = 1),
    nthresh = 3,
    d = 1,
    criterion = "AIC",
    type = "level",
    plot = T,
    trace = T
  )

Setar4.usd <-
  selectSETAR(
    sactnetusd_train,
    m = 4,
    thDelay = seq(0, 3, by = 1),
    nthresh = 3,
    d = 2,
    criterion = "AIC",
    type = "level",
    plot = T,
    trace = T
  )


Setar1.usd$allTh%>%
  as.data.frame()%>%
  arrange(AIC,thDelay,mL,mH)%>%
  head(5)


Setar2.usd$allTh%>%
  as.data.frame()%>%
  arrange(AIC,thDelay,mL,mH)%>%
  head(5)

Setar3.usd$allTh%>%
  as.data.frame()%>%
  arrange(AIC,thDelay,mL,mH)%>%
  head(5)

Setar4.usd$allTh%>%
  as.data.frame()%>%
  arrange(AIC,thDelay,mL,mH)%>%
  head(5)
```

```{r}

modeloas1.usd <-
  setar(
    sactnetusd_train,
    mL = 1,
    mH = 3,
    d=1,
    nthresh = 1,
    thDelay = 2,
    type = "level"
  )
## Raiz Unitaria
summary(modeloas1.usd) #residuals variance = 0.005525,  AIC = -632, MAPE = 0.4352%
# plot(modeloas1)
checkresiduals(ts(modeloas1.usd$residuals,start=c(2011,1),frequency = 12))

modeloas2.usd <-
  setar(
    sactnetusd_train,
    mL = 3,
    mH = 1,
    d=2,
    nthresh = 1,
    thDelay = 1,
    type = "level"
  )
## Raiz Unitaria
summary(modeloas2.usd) # residuals variance = 0.005857,  AIC = -635, MAPE = 0.4584%
# plot(modeloas2)
checkresiduals(ts(modeloas2.usd$residuals,start=c(2011,1),frequency = 12))

modeloas3.usd <-
  setar(
    sactnetusd_train,
    mL = 1,
    mH = 1,
    d=1,
    nthresh = 1,
    thDelay = 1,
    type = "level"
  )
## Raiz Unitaria
summary(modeloas3.usd) # residuals variance = 0.006319,  AIC = -621, MAPE = 0.4621%
# plot(modeloas3)
checkresiduals(ts(modeloas3.usd$residuals,start=c(2011,1),frequency = 12))

modeloas4.usd <-
  setar(
    sactnetusd_train,
    m = 3,
    mL = 1,
    mH = 2,
    d=2,
    nthresh = 1,
    thDelay = 0,
    type = "level"
  )
summary(modeloas4.usd) # residuals variance = 0.006319,  AIC = -621, MAPE = 0.4621%
# plot(modeloas4)
checkresiduals(ts(modeloas4.usd$residuals,start=c(2011,1),frequency = 12))


AIC(modeloas1.usd)
AIC(modeloas2.usd)
AIC(modeloas3.usd)
AIC(modeloas4.usd)
```

```{r}
pronsetar1.usd<- predict(modeloas1.usd, n.ahead = 7)
pronsetar2.usd<- predict(modeloas2.usd, n.ahead = 7)
pronsetar3.usd<- predict(modeloas3.usd, n.ahead = 7)
pronsetar4.usd<- predict(modeloas4.usd, n.ahead = 7)

fit1.usd<-ts(modeloas1.usd$fitted.values,start =c(2011,1),frequency = 12)
fit2.usd<-ts(modeloas2.usd$fitted.values,start =c(2011,1),frequency = 12)
fit3.usd<-ts(modeloas3.usd$fitted.values,start =c(2011,1),frequency = 12)
fit4.usd<-ts(modeloas4.usd$fitted.values,start =c(2011,1),frequency = 12)

autoplot(sactnetusd_train)+
  autolayer(fit1.usd)+
  autolayer(fit2.usd)+
  autolayer(fit3.usd)+
  autolayer(fit4.usd)+
  theme_bw()

data.frame(
Modelo= c(
  "1) m = 3,mL = 3,mH = 1, d=1",
  "2) m = 3,mL = 2,mH = 3, d=2",
  "3) m = 3,mL = 3,mH = 2, d=1",
  "4) m = 3,mL = 1,mH = 2, d=2"
),
RMSE=c(
  Metrics::rmse(sactnetusd_test, pronsetar1.usd),
  Metrics::rmse(sactnetusd_test, pronsetar2.usd),
  Metrics::rmse(sactnetusd_test, pronsetar3.usd),
  Metrics::rmse(sactnetusd_test, pronsetar4.usd)))%>%
  arrange(RMSE)%>%
  knitr::kable()

autoplot(sactnetusd_test)+
  autolayer(pronsetar1.usd)+
  autolayer(pronsetar2.usd)+
  autolayer(pronsetar3.usd)+
  autolayer(pronsetar4.usd)+
  theme_bw()
```




### Metricas Generales

```{r , eval=T}
Metrics::rmse(sactnetusd_test,prontar1.usd)
Metrics::rmse(sactnetusd_test, pronsetar4.usd)


autoplot(sactnetusd_test)+
  autolayer(prontar1.usd)+
  autolayer(pronsetar4.usd)+
  theme_bw()
```
