---
title: "Modeltime: Demo"
author: "Sergio Cubero"
date: "1/4/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

Ejemplo tomado: 

```{r ,echo=FALSE}
vembedr::embed_youtube("6RjYIOCnRMk")
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r paquetes}
# Machine Learning
library(tidymodels)
library(modeltime)
library(modeltime.ensemble)
library(modeltime.resample)

# Time Series
library(timetk)

# Core
library(tidyverse)

# Load data
library(readxl)
```



```{r , include=F}
SeriesDatos <- read_excel("~/Google Drive/Mi unidad/1.Maestria II Ciclo 2021/Curso de Analisis De Casos/Caso II/Datos/Base Datos.xlsx")%>%
  janitor::clean_names()%>%
  mutate(ActivoNeto=paste0(activo_neto,"-01"))%>%
  rename('ActNetCRC'=crc,
         'ActNetUSD'=usd)%>%
  select(ActivoNeto,ActNetCRC,ActNetUSD)
 

colones <- SeriesDatos%>%
  select(Date=ActivoNeto,value=ActNetCRC)%>%
  mutate(Date=as.Date(Date)
         )

dolares <- SeriesDatos%>%
  select(Date=ActivoNeto,value=ActNetUSD)%>%
  mutate(Date=as.Date(Date)
         )
```

# MODELO DE MACHINE LEARNING {.tabset}

## COLONES

```{r}
colones%>%
  plot_time_series(Date,value,.facet_ncol = 3, .interactive = F)
```

### DATA PREPARATION

```{r}
FORECAST_HORIZON <- 5
```

#### Full = Training + Forecast Dataset

```{r}
full_data_tbl <- colones%>%
  select(Date,value)%>%
  future_frame(
    .date_var = Date,
    .length_out = FORECAST_HORIZON,
    .bind_data = T
  )

```

#### Training Data

```{r}
data_prepared_tbl <- full_data_tbl%>%
  filter(!is.na(value))
  
data_prepared_tbl%>%
  tk_summary_diagnostics()
```

#### Future Data Forecast

```{r}
future_tbl <- full_data_tbl%>%
  filter(is.na(value))
```

### SPLITTING

```{r}
splits <- data_prepared_tbl%>%
  arrange(Date)%>%
  time_series_split(
    data_var=Date,
    assess = FORECAST_HORIZON,
    cumulative = T
  )

splits

```

### PREPROCESOR

```{r}
recipe_spec_1 <- recipe(value~., training(splits))%>%
  step_timeseries_signature(Date)%>%
  ## Elimina las columnas o atributos que no aportan
  step_rm(matches("(.iso$)|(.xts)|(day)|(hour)|(minute)|(second)|(am.pm)|(week)")) %>%
  step_normalize(Date_index.num,Date_year)%>%
  step_mutate(Date_month = factor(Date_month,ordered = T))%>%
  step_dummy(all_nominal(),one_hot = T)

recipe_spec_1 %>% prep() %>% juice() %>% glimpse()

recipe_spec_2 <- recipe_spec_1%>%
  update_role(Date,new_role = "ID")

recipe_spec_2 %>% prep() %>% juice() %>% glimpse()

recipe_spec_1 %>% prep() %>% summary()
recipe_spec_2 %>% prep() %>% summary()
```

### MODELS

#### autoarima xgboost 

```{r}
wflw_fit_autoarima_boost <- workflow()%>%
  add_model(
    arima_boost(
    min_n = 2,
    learn_rate = 0.015
) %>%
    set_engine(engine = "auto_arima_xgboost")
  )%>%
  add_recipe(recipe_spec_1)%>%
  fit(training(splits))
```

#### prophet 

```{r}
wflw_fit_prophet <- workflow()%>%
  add_model(
    prophet_reg() %>% set_engine("prophet")
  )%>%
  add_recipe(recipe_spec_1)%>%
  fit(training(splits))
```

#### XGBOOST

```{r}
wflw_fit_xgboost_0_015 <- workflow()%>%
  add_model(
    boost_tree(learn_rate=0.15) %>% set_engine("xgboost")
  )%>%
  add_recipe(recipe_spec_2)%>%
  fit(training(splits))

wflw_fit_xgboost_0_1 <- workflow()%>%
  add_model(
    boost_tree(learn_rate=0.1) %>% set_engine("xgboost")
  )%>%
  add_recipe(recipe_spec_2)%>%
  fit(training(splits))

wflw_fit_xgboost_0_3 <- workflow()%>%
  add_model(
    boost_tree(learn_rate=0.3) %>% set_engine("xgboost")
  )%>%
  add_recipe(recipe_spec_2)%>%
  fit(training(splits))

```

#### Random Forest

```{r}
wflw_fit_rf_1000 <- workflow()%>%
  add_model(
    rand_forest(
                trees = 1000
                ) %>% 
      set_engine("randomForest")
  )%>%
  add_recipe(recipe_spec_2)%>%
  fit(training(splits))

wflw_fit_rf_500 <- workflow()%>%
  add_model(
    rand_forest(
                trees = 500
                ) %>% 
      set_engine("randomForest")
  )%>%
  add_recipe(recipe_spec_2)%>%
  fit(training(splits))

wflw_fit_rf_200 <- workflow()%>%
  add_model(
    rand_forest(
                trees = 200
                ) %>% 
      set_engine("randomForest")
  )%>%
  add_recipe(recipe_spec_2)%>%
  fit(training(splits))
```

#### SVM

```{r}
wflw_fit_svm <- workflow()%>%
  add_model(
    svm_rbf() %>% set_engine("kernlab")
  )%>%
  add_recipe(recipe_spec_2)%>%
  fit(training(splits))
```

#### prophet_boost

```{r}
wflw_fit_prophet_boost <- workflow()%>%
  add_model(
    prophet_boost(
      seasonality_yearly = F,
      seasonality_weekly = F,
      seasonality_daily =  F,
    ) %>% 
      set_engine("prophet_xgboost")
  )%>%
  add_recipe(recipe_spec_1)%>%
  fit(training(splits))
```

### MODELTIME WORKFLOW

#### modeltime table

```{r}
submodels_tbl <- modeltime_table(
  wflw_fit_autoarima_boost,
  #wflw_fit_prophet, #1
  wflw_fit_prophet_boost, #2
  #wflw_fit_xgboost_0_015, #3
  #wflw_fit_xgboost_0_1, #4
  wflw_fit_xgboost_0_3, #5
  #wflw_fit_rf_1000, #6
  wflw_fit_rf_500 #, #7
  #wflw_fit_rf_200, #8
  #wflw_fit_svm #9
)

submodels_tbl
```

#### calibrate Testing Data

```{r}
submodels_calibrated_tbl <- submodels_tbl %>%
  modeltime_calibrate(testing(splits))

submodels_calibrated_tbl
```

#### Measure Test Accuracy

```{r}
submodels_calibrated_tbl%>% 
  modeltime_accuracy()%>%
  arrange(rmse)
```

#### Visualize test forecast

```{r}
submodels_calibrated_tbl %>%
  modeltime_forecast(
    new_data = testing(splits),
    actual_data = data_prepared_tbl,
    keep_data = T
  )%>%
  plot_modeltime_forecast(
    .facet_ncol=2
  )
```

#### Refit on full training dataset

```{r}
submodels_refit_tbl <- submodels_calibrated_tbl %>%
  modeltime_refit(data_prepared_tbl)
```

#### Visualize Submodel Forecast

```{r}
submodels_refit_tbl%>%
  modeltime_forecast(
    new_data =  future_tbl,
    actual_data = data_prepared_tbl,
    keep_data = T
  )%>%
  plot_modeltime_forecast(
    .facet_ncol=2
  )
```

#### Validación cruzada

https://cran.r-project.org/web/packages/modeltime.resample/vignettes/getting-started.html

```{r}
resamples_tscv <- time_series_cv(
    data        = data_prepared_tbl,
    date_var    = Date,
    assess      = FORECAST_HORIZON,
    initial     = "36 month",
    skip        = FORECAST_HORIZON,
    slice_limit = 5
)

resamples_tscv

resamples_tscv %>%
    tk_time_series_cv_plan() %>%
    plot_time_series_cv_plan(Date, 
                             value, 
                             .facet_ncol = 2,
                             .interactive = T)
```

##### Generate Resample Predictions

```{r}
resamples_fitted <- submodels_tbl %>%
    modeltime_fit_resamples(
        resamples = resamples_tscv,
        control   = control_resamples(verbose = FALSE)
    )

resamples_fitted
```

##### Evaluate the Results

```{r}
resamples_fitted %>%
    plot_modeltime_resamples(
      .point_size  = 3, 
      .point_alpha = 0.8,
      .interactive = T
    )

resamples_fitted %>%
    modeltime_resample_accuracy(summary_fns = mean) %>%
    table_modeltime_accuracy(.interactive = T)
```

### ENSEMBLE

#### Ensamble Media y Meta-Learner

```{r}
ensemble_fit_mean <- submodels_tbl %>%
  #filter(!.model_id %in% c(1))%>%
  ensemble_average(type="mean")


ensemble_fit_lm <- resamples_fitted %>%
  ensemble_model_spec(
    model_spec = linear_reg(
      penalty = tune(),
      mixture = tune()
    ) %>%
      set_engine("glmnet"),
    grid = 2,
    control = control_grid(verbose = TRUE)
  )

ensemble_fit_xg<- resamples_fitted %>%
  ensemble_model_spec(
    model_spec = boost_tree(
      mtry=tune(),
      trees=tune(),
      learn_rate = tune()
    ) %>% set_engine("xgboost"),
    control = control_grid(verbose = TRUE)
  )
```

```{r}
ensemble_tbl<- modeltime_table(
  ensemble_fit_mean,
  ensemble_fit_lm,
  ensemble_fit_xg
)
```


#### Ensemble test Accuracy

```{r}
ensemble_tbl%>%
  combine_modeltime_tables(submodels_tbl)%>%
  modeltime_accuracy(testing(splits))%>%
  arrange(rmse,mae,mape,mase)%>%
  knitr::kable()
```

#### Ensemble Test Forecast

```{r}
ensemble_tbl%>%
  modeltime_calibrate(testing(splits))%>%
  modeltime_forecast(
    new_data =  testing(splits),
    actual_data = data_prepared_tbl,
    keep_data = T,
    conf_by_id = T,
    conf_interval = 0.95
  )%>%
  plot_modeltime_forecast(
    .facet_ncol=2
  )

ensemble_tbl%>%
  combine_modeltime_tables(submodels_tbl)%>%
  modeltime_calibrate(testing(splits))%>%
  modeltime_forecast(
    new_data =  testing(splits),
    actual_data = data_prepared_tbl,
    keep_data = T
  )%>%
  plot_modeltime_forecast(
    .facet_ncol=2
  )

ensemble_tbl_all_model<-ensemble_tbl%>%
  combine_modeltime_tables(submodels_tbl)
```

#### Refit Ensemble

```{r}
ensemble_refit_tbl <- ensemble_tbl%>%
  modeltime_refit(data_prepared_tbl)
```

#### Visualize Ensemble Forecast

```{r}
ensemble_refit_tbl%>%
  modeltime_forecast(
    new_data =  future_tbl,
    actual_data = data_prepared_tbl,
    keep_data = T
  )%>%
  plot_modeltime_forecast(
    .facet_ncol=2
  )

```



## DOLARES

### DATA

```{r}
dolares%>%
  plot_time_series(Date,value,.facet_ncol = 3, .interactive = F)
```

### DATA PREPARATION

```{r}
FORECAST_HORIZON <- 5
```

#### Full = Training + Forecast Dataset

```{r}
full_data_tbl <- dolares%>%
  select(Date,value)%>%
  future_frame(
    .date_var = Date,
    .length_out = FORECAST_HORIZON,
    .bind_data = T
  )

```

#### Training Data

```{r}
data_prepared_tbl <- full_data_tbl%>%
  filter(!is.na(value))
  
data_prepared_tbl%>%
  tk_summary_diagnostics()
```

#### Future Data Forecast

```{r}
future_tbl <- full_data_tbl%>%
  filter(is.na(value))
```

### SPLITTING

```{r}
splits <- data_prepared_tbl%>%
  arrange(Date)%>%
  time_series_split(
    data_var=Date,
    assess = FORECAST_HORIZON,
    cumulative = T
  )

splits

```

### PREPROCESOR

```{r}
recipe_spec_1 <- recipe(value~., training(splits))%>%
  step_timeseries_signature(Date)%>%
  ## Elimina las columnas o atributos que no aportan
  step_rm(matches("(.iso$)|(.xts)|(day)|(hour)|(minute)|(second)|(am.pm)|(week)")) %>%
  step_normalize(Date_index.num,Date_year)%>%
  step_mutate(Date_month = factor(Date_month,ordered = T))%>%
  step_dummy(all_nominal(),one_hot = T)

recipe_spec_1 %>% prep() %>% juice() %>% glimpse()

recipe_spec_2 <- recipe_spec_1%>%
  update_role(Date,new_role = "ID")

recipe_spec_2 %>% prep() %>% juice() %>% glimpse()

recipe_spec_1 %>% prep() %>% summary()
recipe_spec_2 %>% prep() %>% summary()
```

### MODELS

#### autoarima xgboost 

```{r}
wflw_fit_autoarima_boost <- workflow()%>%
  add_model(
    arima_boost(
    min_n = 2,
    learn_rate = 0.015
) %>%
    set_engine(engine = "auto_arima_xgboost")
  )%>%
  add_recipe(recipe_spec_1)%>%
  fit(training(splits))
```

#### prophet 

```{r}
wflw_fit_prophet <- workflow()%>%
  add_model(
    prophet_reg() %>% set_engine("prophet")
  )%>%
  add_recipe(recipe_spec_1)%>%
  fit(training(splits))
```

#### XGBOOST

```{r}
wflw_fit_xgboost_0_015 <- workflow()%>%
  add_model(
    boost_tree(learn_rate=0.15) %>% set_engine("xgboost")
  )%>%
  add_recipe(recipe_spec_2)%>%
  fit(training(splits))

wflw_fit_xgboost_0_1 <- workflow()%>%
  add_model(
    boost_tree(learn_rate=0.1) %>% set_engine("xgboost")
  )%>%
  add_recipe(recipe_spec_2)%>%
  fit(training(splits))

wflw_fit_xgboost_0_3 <- workflow()%>%
  add_model(
    boost_tree(learn_rate=0.3) %>% set_engine("xgboost")
  )%>%
  add_recipe(recipe_spec_2)%>%
  fit(training(splits))

```

#### Random Forest

```{r}
wflw_fit_rf_1000 <- workflow()%>%
  add_model(
    rand_forest(
                trees = 1000
                ) %>% 
      set_engine("randomForest")
  )%>%
  add_recipe(recipe_spec_2)%>%
  fit(training(splits))

wflw_fit_rf_500 <- workflow()%>%
  add_model(
    rand_forest(
                trees = 500
                ) %>% 
      set_engine("randomForest")
  )%>%
  add_recipe(recipe_spec_2)%>%
  fit(training(splits))

wflw_fit_rf_200 <- workflow()%>%
  add_model(
    rand_forest(
                trees = 200
                ) %>% 
      set_engine("randomForest")
  )%>%
  add_recipe(recipe_spec_2)%>%
  fit(training(splits))
```

#### SVM

```{r}
wflw_fit_svm <- workflow()%>%
  add_model(
    svm_rbf() %>% set_engine("kernlab")
  )%>%
  add_recipe(recipe_spec_2)%>%
  fit(training(splits))
```

#### prophet_boost

```{r}
wflw_fit_prophet_boost <- workflow()%>%
  add_model(
    prophet_boost(
      seasonality_yearly = F,
      seasonality_weekly = F,
      seasonality_daily =  F,
    ) %>% 
      set_engine("prophet_xgboost")
  )%>%
  add_recipe(recipe_spec_1)%>%
  fit(training(splits))
```

### MODELTIME WORKFLOW

#### modeltime table

```{r}
submodels_tbl <- modeltime_table(
  wflw_fit_autoarima_boost,
  #wflw_fit_prophet, #1
  wflw_fit_prophet_boost, #2
  #wflw_fit_xgboost_0_015, #3
  #wflw_fit_xgboost_0_1, #4
  wflw_fit_xgboost_0_3, #5
  #wflw_fit_rf_1000, #6
  wflw_fit_rf_500 #, #7
  #wflw_fit_rf_200, #8
  #wflw_fit_svm #9
)

submodels_tbl
```

#### calibrate Testing Data

```{r}
submodels_calibrated_tbl <- submodels_tbl %>%
  modeltime_calibrate(testing(splits))

submodels_calibrated_tbl
```

#### Measure Test Accuracy

```{r}
submodels_calibrated_tbl%>% 
  modeltime_accuracy()%>%
  arrange(rmse)
```

#### Visualize test forecast

```{r}
submodels_calibrated_tbl %>%
  modeltime_forecast(
    new_data = testing(splits),
    actual_data = data_prepared_tbl,
    keep_data = T
  )%>%
  plot_modeltime_forecast(
    .facet_ncol=2
  )
```

#### Refit on full training dataset

```{r}
submodels_refit_tbl <- submodels_calibrated_tbl %>%
  modeltime_refit(data_prepared_tbl)
```

#### Visualize Submodel Forecast

```{r}
submodels_refit_tbl%>%
  modeltime_forecast(
    new_data =  future_tbl,
    actual_data = data_prepared_tbl,
    keep_data = T
  )%>%
  plot_modeltime_forecast(
    .facet_ncol=2
  )
```

#### Validación cruzada

https://cran.r-project.org/web/packages/modeltime.resample/vignettes/getting-started.html

```{r}
resamples_tscv <- time_series_cv(
    data        = data_prepared_tbl,
    date_var    = Date,
    assess      = FORECAST_HORIZON,
    initial     = "36 month",
    skip        = FORECAST_HORIZON,
    slice_limit = 5
)

resamples_tscv

resamples_tscv %>%
    tk_time_series_cv_plan() %>%
    plot_time_series_cv_plan(Date, 
                             value, 
                             .facet_ncol = 2,
                             .interactive = T)
```

##### Generate Resample Predictions

```{r}
resamples_fitted <- submodels_tbl %>%
    modeltime_fit_resamples(
        resamples = resamples_tscv,
        control   = control_resamples(verbose = FALSE)
    )

resamples_fitted
```

##### Evaluate the Results

```{r}
resamples_fitted %>%
    plot_modeltime_resamples(
      .point_size  = 3, 
      .point_alpha = 0.8,
      .interactive = T
    )

resamples_fitted %>%
    modeltime_resample_accuracy(summary_fns = mean) %>%
    table_modeltime_accuracy(.interactive = T)
```

### ENSEMBLE

#### Ensamble Media y Meta-Learner

```{r}
ensemble_fit_mean <- submodels_tbl %>%
  #filter(!.model_id %in% c(1))%>%
  ensemble_average(type="mean")


ensemble_fit_lm <- resamples_fitted %>%
  ensemble_model_spec(
    model_spec = linear_reg(
      penalty = tune(),
      mixture = tune()
    ) %>%
      set_engine("glmnet"),
    grid = 2,
    control = control_grid(verbose = TRUE)
  )

ensemble_fit_xg<- resamples_fitted %>%
  ensemble_model_spec(
    model_spec = boost_tree(
      mtry=tune(),
      trees=tune(),
      learn_rate = tune()
    ) %>% set_engine("xgboost"),
    control = control_grid(verbose = TRUE)
  )
```

```{r}
ensemble_tbl<- modeltime_table(
  ensemble_fit_mean,
  ensemble_fit_lm,
  ensemble_fit_xg
)
```


#### Ensemble test Accuracy

```{r}
ensemble_tbl%>%
  combine_modeltime_tables(submodels_tbl)%>%
  modeltime_accuracy(testing(splits))%>%
  arrange(rmse,mae,mape,mase)%>%
  knitr::kable()
```

#### Ensemble Test Forecast

```{r}
ensemble_tbl%>%
  modeltime_calibrate(testing(splits))%>%
  modeltime_forecast(
    new_data =  testing(splits),
    actual_data = data_prepared_tbl,
    keep_data = T,
    conf_by_id = T,
    conf_interval = 0.95
  )%>%
  plot_modeltime_forecast(
    .facet_ncol=2
  )

ensemble_tbl%>%
  combine_modeltime_tables(submodels_tbl)%>%
  modeltime_calibrate(testing(splits))%>%
  modeltime_forecast(
    new_data =  testing(splits),
    actual_data = data_prepared_tbl,
    keep_data = T
  )%>%
  plot_modeltime_forecast(
    .facet_ncol=2
  )

ensemble_tbl_all_model<-ensemble_tbl%>%
  combine_modeltime_tables(submodels_tbl)
```

#### Refit Ensemble

```{r}
ensemble_refit_tbl <- ensemble_tbl%>%
  modeltime_refit(data_prepared_tbl)
```

#### Visualize Ensemble Forecast

```{r}
ensemble_refit_tbl%>%
  modeltime_forecast(
    new_data =  future_tbl,
    actual_data = data_prepared_tbl,
    keep_data = T
  )%>%
  plot_modeltime_forecast(
    .facet_ncol=2
  )

```

