---
title: "No lineales"
author: "Sergio Cubero"
date: "9/17/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(forecast)
library(TSA)
library(NTS)
library(MSwM)
library(tsDyn)
library(fNonlinear)
library(dlm)
library(astsa)
library(tidyverse)
library(readxl)
```


```{r datos, include=FALSE}

SeriesDatos <- read_excel("~/Google Drive/Mi unidad/1.Maestria II Ciclo 2021/Curso de Analisis De Casos/Caso II/Datos/Base Datos.xlsx")%>%
  janitor::clean_names()%>%
  mutate(ActivoNeto=paste0(activo_neto,"-01"))%>%
  rename('ActNetCRC'=crc,
         'ActNetUSD'=usd)

actnetcrc<- ts(SeriesDatos[,2],start =c(2001,2),end=c(2021,7), frequency = 12)
actnetusd<- ts(SeriesDatos[,3],start =c(2001,2),end=c(2021,7), frequency = 12)
actnet <- cbind(actnetcrc,actnetusd) 

## Peridos de Tiempo
inicio_train<- c(2011,1)
fin_train<- c(2020,12)

inicio_test <- c(2021,1)

sactnetcrc<- window(actnetcrc,start=inicio_train)
sactnetcrc_train<- window(actnetcrc,start=inicio_train, end=fin_train)
sactnetcrc_test<- window(actnetcrc,start=inicio_test)

sactnetusd<- window(actnetusd,start=inicio_train)
sactnetusd_train<- window(actnetusd,start=inicio_train, end=fin_train)
sactnetusd_test<- window(actnetusd,start=inicio_test)
```



# Pruebas de No Linealidad Colones

## Colones

### General

```{r}
nonlinearTseries::nonlinearityTest(actnetcrc, verbose = TRUE)
```

### Teraesvirta Neural Network Test
```{r}
fNonlinear::tnnTest(actnetcrc, lag = 1, title = NULL, description = NULL)
```

### Kennan tests for nonlineary

La hipótesis nula de que la serie de tiempo sigue algún proceso de AR.

```{r}
Keenan.test(actnetcrc)
Keenan.test(actnetcrc, order=1)
Keenan.test(actnetcrc, order=2)
Keenan.test(actnetcrc, order=3)
```


### BDS

```{r}
bdsTest(actnetcrc)
```


### Thresold Non Linearity

```{r}
lag1.plot(actnetcrc, max.lag=12)
thr.test(actnetcrc)
thr.test(actnetcrc,p=2,d=1)
thr.test(actnetcrc,p=2,d=2)
thr.test(actnetcrc,p=3,d=1)
thr.test(actnetcrc,p=3,d=2)
thr.test(actnetcrc,p=3,d=3)
```

## Dolares

### General

```{r}
nonlinearTseries::nonlinearityTest(actnetusd, verbose = TRUE)
```


### Teraesvirta Neural Network Test
```{r}
fNonlinear::tnnTest(actnetusd, lag = 1, title = NULL, description = NULL)
```

### Kennan tests for nonlineary

La hipótesis nula de que la serie de tiempo sigue algún proceso de AR.

```{r}
Keenan.test(actnetusd)
Keenan.test(actnetusd, order=1)
Keenan.test(actnetusd, order=2)
Keenan.test(actnetusd, order=3)
```


### BDS

```{r}
bdsTest(actnetusd)
```


### Thresold Non Linearity

```{r}
lag1.plot(actnetusd, max.lag=12)
thr.test(actnetusd)
thr.test(actnetusd,p=2,d=1)
thr.test(actnetusd,p=2,d=2)
thr.test(actnetusd,p=3,d=1)
thr.test(actnetusd,p=3,d=2)
thr.test(actnetusd,p=3,d=3)
```



# Modelos No Lineales

## Colones

### TAR

```{r ,eval=F}

set.seed(777)
# m orden
pm <- 1:4

mod.list.tar<-list()



for(j in pm){
  for(i in pm){
    for(l in pm){
      mod<-TSA::tar(sactnetcrc_train,p1=j,p2=i,d=l,print=TRUE)  
      
      mod.list.tar[[paste(j,l,i,sep="-")]]<-mod$AIC
      print(paste(j,l,i,sep="-"))    
    }
  }
}

AICTar<-bind_rows(mod.list.tar,.id = "Ordene-delay")%>%
  arrange(`1`)

knitr::kable(head(AICTar,20))

```

```{r ,eval=F}

# m orden
pm <- 1:3

mod.list.tar<-list()
AIC.best.list<-list()

AICM = NULL
model.best <- list(d=0, p1=0, p2=0)
AIC.best = 2888

for(l in pm){
  for(j in pm){
    for(i in pm){
      set.seed(777)
      model.tar.s = tar(sactnetcrc_train,p1=j,p2=i,d=l)
      mod.list.tar[[paste(j,i,l,sep="-")]]<-model.tar.s$AIC
      #print(paste(j,i,l,sep="-"))    
      
      if (model.tar.s$AIC < AIC.best) {
		    AIC.best = model.tar.s$AIC
		    AIC.best.list[[paste(j,i,l,sep="-")]]<-AIC.best
		    #print(AIC.best)
		    model.best$d = l
		    model.best$p1 = model.tar.s$p1
		    model.best$p2 = model.tar.s$p2 
		    print(paste(model.tar.s$p1,model.tar.s$p2,l,sep="-")) }
    }
  }
}

AICTar<-bind_rows(mod.list.tar,.id = "Ordene-delay")%>%
  arrange(`1`)

knitr::kable(head(AICTar,20))

AICTarBest<-bind_rows(AIC.best.list,.id = "Ordene-delay")%>%
  arrange(`1`)

knitr::kable(head(AICTarBest,20))

```

```{r}
mod.tar1<-TSA::tar(sactnetcrc_train,p1=2,p2=3,d=1)  
mod.tar2<-TSA::tar(sactnetcrc_train,p1=3,p2=1,d=1)  
mod.tar3<-TSA::tar(sactnetcrc_train,p1=3,p2=2,d=1)  

mod.tar1$thd
mod.tar2$thd
mod.tar3$thd

mod.tar1$qr1$coefficients
mod.tar2$qr1$coefficients
mod.tar3$qr1$coefficients

mod.tar1$qr2$coefficients
mod.tar2$qr2$coefficients
mod.tar3$qr2$coefficients

mod.tar1$AIC
mod.tar2$AIC
mod.tar3$AIC

#tsdiag(mod.tar1)
tsdiag(mod.tar2)
#tsdiag(mod.tar3)


checkresiduals(ts(mod.tar1$residuals,start=c(2011,1),frequency = 12))
checkresiduals(ts(mod.tar2$residuals,start=c(2011,1),frequency = 12))
checkresiduals(ts(mod.tar3$residuals,start=c(2011,1),frequency = 12))
```


```{r}
prontar1<- ts(as.vector(predict(mod.tar1,n.ahead=7,n.sim=1000)$fit),start=c(2021,1),frequency = 12)
prontar2<- ts(as.vector(predict(mod.tar2,n.ahead=7,n.sim=1000)$fit),start=c(2021,1),frequency = 12)
prontar3<- ts(as.vector(predict(mod.tar3,n.ahead=7,n.sim=1000)$fit),start=c(2021,1),frequency = 12)

fit1<-ts(as.vector(mod.tar1$y)-as.vector(mod.tar1$residuals),start =c(2011,1),frequency = 12)
fit2<-ts(sactnetcrc_train-mod.tar2$residuals,start =c(2011,1),frequency = 12)
fit3<-ts(sactnetcrc_train-mod.tar3$residuals,start =c(2011,1),frequency = 12)

autoplot(sactnetcrc_train)+
  autolayer(fit1)+
  autolayer(fit2)+
  autolayer(fit3)+
  theme_bw()

Metrics::rmse(sactnetcrc_test, prontar1)
Metrics::rmse(sactnetcrc_test, prontar2)
Metrics::rmse(sactnetcrc_test, prontar3)

autoplot(sactnetcrc_test)+
  autolayer(prontar1)+
  autolayer(prontar2)+
  autolayer(prontar3)+
  theme_bw()
```

### SETAR

Thus the threshold delay, the number of lags in each regime and the threshold value are computed.

```{r, eval=F}

Setar1 <-
  selectSETAR(
    sactnetcrc_train, 
    include = c("const", "trend","none", "both"),
    m = 3,
    thDelay = seq(1, 2, by = 1),
    nthresh = 2,
    criterion = "AIC",
    type = "level",
    plot = T,
    trace = T
  )

Setar2 <-
  selectSETAR(
    sactnetcrc_train,
    m = 3,
    d=2,
    thDelay = seq(1, 2, by = 1),
    nthresh = 2,
    criterion = "AIC",
    type = "level",
    plot = T,
    trace = T
  )

Setar3 <-
  selectSETAR(
    sactnetcrc_train,
    m = 3,
    thDelay = seq(0, 2, by = 1),
    nthresh = 1,
    d = 1,
    criterion = "AIC",
    type = "level",
    plot = T,
    trace = T
  )

Setar4 <-
  selectSETAR(
    sactnetcrc_train,
    m = 3,
    thDelay = seq(0, 2, by = 1),
    nthresh = 1,
    d = 2,
    criterion = "AIC",
    type = "level",
    plot = T,
    trace = T
  )


Setar1$allTh%>%
  as.data.frame()%>%
  arrange(AIC,thDelay,mL,mH)%>%
  head(5)


Setar2$allTh%>%
  as.data.frame()%>%
  arrange(AIC,thDelay,mL,mH)%>%
  head(5)

Setar3$allTh%>%
  as.data.frame()%>%
  arrange(AIC,thDelay,mL,mH)%>%
  head(5)

Setar4$allTh%>%
  as.data.frame()%>%
  arrange(AIC,thDelay,mL,mH)%>%
  head(5)
```

```{r}

modeloas1 <-
  setar(
    sactnetcrc_train,
    m = 3,
    mL = 3,
    mH = 1,
    d=1,
    nthresh = 1,
    thDelay = 2,
    type = "level"
  )
## Raiz Unitaria
summary(modeloas1) #residuals variance = 0.005525,  AIC = -632, MAPE = 0.4352%
# plot(modeloas1)
checkresiduals(ts(modeloas1$residuals,start=c(2011,1),frequency = 12))

modeloas2 <-
  setar(
    sactnetcrc_train,
    m = 3,
    mL = 2,
    mH = 3,
    d=2,
    nthresh = 1,
    thDelay = 2,
    type = "level"
  )
## Raiz Unitaria
summary(modeloas2) # residuals variance = 0.005857,  AIC = -635, MAPE = 0.4584%
# plot(modeloas2)
checkresiduals(ts(modeloas2$residuals,start=c(2011,1),frequency = 12))

modeloas3 <-
  setar(
    sactnetcrc_train,
    m = 3,
    mL = 3,
    mH = 2,
    d=1,
    nthresh = 1,
    thDelay = 0,
    type = "level"
  )
## Raiz Unitaria
summary(modeloas3) # residuals variance = 0.006319,  AIC = -621, MAPE = 0.4621%
# plot(modeloas3)
checkresiduals(ts(modeloas3$residuals,start=c(2011,1),frequency = 12))

modeloas4 <-
  setar(
    sactnetcrc_train,
    m = 3,
    mL = 1,
    mH = 2,
    d=2,
    nthresh = 1,
    thDelay = 0,
    type = "level"
  )
summary(modeloas4) # residuals variance = 0.006319,  AIC = -621, MAPE = 0.4621%
# plot(modeloas4)
checkresiduals(ts(modeloas4$residuals,start=c(2011,1),frequency = 12))


AIC(modeloas1)
AIC(modeloas2)
AIC(modeloas3)
AIC(modeloas4)
```

```{r}
pronsetar1<- predict(modeloas1, n.ahead = 7)
pronsetar2<- predict(modeloas2, n.ahead = 7)
pronsetar3<- predict(modeloas3, n.ahead = 7)
pronsetar4<- predict(modeloas4, n.ahead = 7)

fit1<-ts(modeloas1$fitted.values,start =c(2011,1),frequency = 12)
fit2<-ts(modeloas2$fitted.values,start =c(2011,1),frequency = 12)
fit3<-ts(modeloas3$fitted.values,start =c(2011,1),frequency = 12)
fit4<-ts(modeloas4$fitted.values,start =c(2011,1),frequency = 12)

autoplot(sactnetcrc_train)+
  autolayer(fit1)+
  autolayer(fit2)+
  autolayer(fit3)+
  autolayer(fit4)+
  theme_bw()

Metrics::rmse(sactnetcrc_test, pronsetar1)
Metrics::rmse(sactnetcrc_test, pronsetar2)
Metrics::rmse(sactnetcrc_test, pronsetar3)
Metrics::rmse(sactnetcrc_test, pronsetar4)

autoplot(sactnetcrc_test)+
  autolayer(exp(pronsetar1))+
  autolayer(exp(pronsetar2))+
  autolayer(exp(pronsetar3))+
  autolayer(exp(pronsetar4))+
  theme_bw()
```


### LSTAR

```{r , eval=F}

modlstat1<-lstar(sactnetcrc_train, m=1, include = c("const", "trend","none", "both"))     
modlstat2<-lstar(sactnetcrc_train, m=2, include = c("const", "trend","none", "both"))  
modlstat3<-lstar(sactnetcrc_train, m=3, include = c("const", "trend","none", "both"))  

plot(modlstat1)
plot(modlstat2)
plot(modlstat3)

summary(modlstat1)
summary(modlstat2)
summary(modlstat3)

pronlsetar1<- predict(modlstat1, n.ahead = 7)
pronlsetar2<- predict(modlstat2, n.ahead = 7)
pronlsetar3<- predict(modlstat3, n.ahead = 7)

Metrics::rmse(sactnetcrc_test, pronlsetar1)
Metrics::rmse(sactnetcrc_test, pronlsetar2)
Metrics::rmse(sactnetcrc_test, pronlsetar3)

fit1<-ts(modlstat1$fitted.values,start =c(2011,1),frequency = 12)
fit2<-ts(modlstat2$fitted.values,start =c(2011,1),frequency = 12)
fit3<-ts(modlstat3$fitted.values,start =c(2011,1),frequency = 12)


autoplot(sactnetcrc_train)+
  autolayer(fit1)+
  autolayer(fit2)+
  autolayer(fit3)+
  theme_bw()


autoplot(sactnetcrc_test)+
  autolayer(pronlsetar1)+
  autolayer(pronlsetar2)+
  autolayer(pronlsetar3)+
  theme_bw()

```

### Metricas Generales

```{r , eval=F}
Metrics::rmse(sactnetcrc_test,(prontar3))
Metrics::rmse(sactnetcrc_test, (pronsetar2))
Metrics::rmse(sactnetcrc_test, (pronlsetar3))

autoplot(sactnetcrc_test)+
  autolayer((prontar3))+
  autolayer((pronsetar2))+
  autolayer((pronlsetar3))+
  theme_bw()
```

## Dolares

### TAR

```{r ,eval=F}

set.seed(777)
# m orden
pm <- 1:4

mod.list.tar<-list()



for(j in pm){
  for(i in pm){
    for(l in pm){
      mod<-TSA::tar(sactnetusd_train,p1=j,p2=i,d=l,print=TRUE)  
      
      mod.list.tar[[paste(j,l,i,sep="-")]]<-mod$AIC
      print(paste(j,l,i,sep="-"))    
    }
  }
}

AICTar<-bind_rows(mod.list.tar,.id = "Ordene-delay")%>%
  arrange(`1`)

knitr::kable(head(AICTar,20))

```

```{r ,eval=F}

# m orden
pm <- 1:3

mod.list.tar<-list()
AIC.best.list<-list()

AICM = NULL
model.best <- list(d=0, p1=0, p2=0)
AIC.best = 2888

for(l in pm){
  for(j in pm){
    for(i in pm){
      set.seed(777)
      model.tar.s = tar(sactnetusd_train,p1=j,p2=i,d=l)
      mod.list.tar[[paste(j,i,l,sep="-")]]<-model.tar.s$AIC
      #print(paste(j,i,l,sep="-"))    
      
      if (model.tar.s$AIC < AIC.best) {
		    AIC.best = model.tar.s$AIC
		    AIC.best.list[[paste(j,i,l,sep="-")]]<-AIC.best
		    #print(AIC.best)
		    model.best$d = l
		    model.best$p1 = model.tar.s$p1
		    model.best$p2 = model.tar.s$p2 
		    print(paste(model.tar.s$p1,model.tar.s$p2,l,sep="-")) }
    }
  }
}

AICTar<-bind_rows(mod.list.tar,.id = "Ordene-delay")%>%
  arrange(`1`)

knitr::kable(head(AICTar,20))

AICTarBest<-bind_rows(AIC.best.list,.id = "Ordene-delay")%>%
  arrange(`1`)

knitr::kable(head(AICTarBest,20))

```

```{r}
mod.tar1<-TSA::tar(sactnetusd_train,p1=2,p2=3,d=1)  
mod.tar2<-TSA::tar(sactnetusd_train,p1=3,p2=1,d=1)  
mod.tar3<-TSA::tar(sactnetusd_train,p1=3,p2=2,d=1)  

mod.tar1$thd
mod.tar2$thd
mod.tar3$thd

mod.tar1$qr1$coefficients
mod.tar2$qr1$coefficients
mod.tar3$qr1$coefficients

mod.tar1$qr2$coefficients
mod.tar2$qr2$coefficients
mod.tar3$qr2$coefficients

mod.tar1$AIC
mod.tar2$AIC
mod.tar3$AIC

#tsdiag(mod.tar1)
tsdiag(mod.tar2)
#tsdiag(mod.tar3)


checkresiduals(ts(mod.tar1$residuals,start=c(2011,1),frequency = 12))
checkresiduals(ts(mod.tar2$residuals,start=c(2011,1),frequency = 12))
checkresiduals(ts(mod.tar3$residuals,start=c(2011,1),frequency = 12))
```


```{r}
prontar1<- ts(as.vector(predict(mod.tar1,n.ahead=7,n.sim=1000)$fit),start=c(2021,1),frequency = 12)
prontar2<- ts(as.vector(predict(mod.tar2,n.ahead=7,n.sim=1000)$fit),start=c(2021,1),frequency = 12)
prontar3<- ts(as.vector(predict(mod.tar3,n.ahead=7,n.sim=1000)$fit),start=c(2021,1),frequency = 12)

fit1<-ts(as.vector(mod.tar1$y)-as.vector(mod.tar1$residuals),start =c(2011,1),frequency = 12)
fit2<-ts(sactnetusd_train-mod.tar2$residuals,start =c(2011,1),frequency = 12)
fit3<-ts(sactnetusd_train-mod.tar3$residuals,start =c(2011,1),frequency = 12)

autoplot(sactnetusd_train)+
  autolayer(fit1)+
  autolayer(fit2)+
  autolayer(fit3)+
  theme_bw()

Metrics::rmse(sactnetcrc_test, prontar1)
Metrics::rmse(sactnetcrc_test, prontar2)
Metrics::rmse(sactnetcrc_test, prontar3)

autoplot(sactnetcrc_test)+
  autolayer(prontar1)+
  autolayer(prontar2)+
  autolayer(prontar3)+
  theme_bw()
```

### SETAR

Thus the threshold delay, the number of lags in each regime and the threshold value are computed.

```{r, eval=F}

Setar1 <-
  selectSETAR(
    sactnetusd_train, 
    include = c("const", "trend","none", "both"),
    m = 3,
    thDelay = seq(1, 2, by = 1),
    nthresh = 2,
    criterion = "AIC",
    type = "level",
    plot = T,
    trace = T
  )

Setar2 <-
  selectSETAR(
    sactnetusd_train,
    m = 3,
    d=2,
    thDelay = seq(1, 2, by = 1),
    nthresh = 2,
    criterion = "AIC",
    type = "level",
    plot = T,
    trace = T
  )

Setar3 <-
  selectSETAR(
    sactnetusd_train,
    m = 3,
    thDelay = seq(0, 2, by = 1),
    nthresh = 1,
    d = 1,
    criterion = "AIC",
    type = "level",
    plot = T,
    trace = T
  )

Setar4 <-
  selectSETAR(
    sactnetusd_train,
    m = 3,
    thDelay = seq(0, 2, by = 1),
    nthresh = 1,
    d = 2,
    criterion = "AIC",
    type = "level",
    plot = T,
    trace = T
  )


Setar1$allTh%>%
  as.data.frame()%>%
  arrange(AIC,thDelay,mL,mH)%>%
  head(5)


Setar2$allTh%>%
  as.data.frame()%>%
  arrange(AIC,thDelay,mL,mH)%>%
  head(5)

Setar3$allTh%>%
  as.data.frame()%>%
  arrange(AIC,thDelay,mL,mH)%>%
  head(5)

Setar4$allTh%>%
  as.data.frame()%>%
  arrange(AIC,thDelay,mL,mH)%>%
  head(5)
```

```{r}

modeloas1 <-
  setar(
    sactnetusd_train,
    m = 3,
    mL = 3,
    mH = 1,
    d=1,
    nthresh = 1,
    thDelay = 2,
    type = "level"
  )
## Raiz Unitaria
summary(modeloas1) #residuals variance = 0.005525,  AIC = -632, MAPE = 0.4352%
# plot(modeloas1)
checkresiduals(ts(modeloas1$residuals,start=c(2011,1),frequency = 12))

modeloas2 <-
  setar(
    sactnetusd_train,
    m = 3,
    mL = 2,
    mH = 3,
    d=2,
    nthresh = 1,
    thDelay = 2,
    type = "level"
  )
## Raiz Unitaria
summary(modeloas2) # residuals variance = 0.005857,  AIC = -635, MAPE = 0.4584%
# plot(modeloas2)
checkresiduals(ts(modeloas2$residuals,start=c(2011,1),frequency = 12))

modeloas3 <-
  setar(
    sactnetusd_train,
    m = 3,
    mL = 3,
    mH = 2,
    d=1,
    nthresh = 1,
    thDelay = 0,
    type = "level"
  )
## Raiz Unitaria
summary(modeloas3) # residuals variance = 0.006319,  AIC = -621, MAPE = 0.4621%
# plot(modeloas3)
checkresiduals(ts(modeloas3$residuals,start=c(2011,1),frequency = 12))

modeloas4 <-
  setar(
    sactnetusd_train,
    m = 3,
    mL = 1,
    mH = 2,
    d=2,
    nthresh = 1,
    thDelay = 0,
    type = "level"
  )
summary(modeloas4) # residuals variance = 0.006319,  AIC = -621, MAPE = 0.4621%
# plot(modeloas4)
checkresiduals(ts(modeloas4$residuals,start=c(2011,1),frequency = 12))


AIC(modeloas1)
AIC(modeloas2)
AIC(modeloas3)
AIC(modeloas4)
```

```{r}
pronsetar1<- predict(modeloas1, n.ahead = 7)
pronsetar2<- predict(modeloas2, n.ahead = 7)
pronsetar3<- predict(modeloas3, n.ahead = 7)
pronsetar4<- predict(modeloas4, n.ahead = 7)

fit1<-ts(modeloas1$fitted.values,start =c(2011,1),frequency = 12)
fit2<-ts(modeloas2$fitted.values,start =c(2011,1),frequency = 12)
fit3<-ts(modeloas3$fitted.values,start =c(2011,1),frequency = 12)
fit4<-ts(modeloas4$fitted.values,start =c(2011,1),frequency = 12)

autoplot(sactnetusd_train)+
  autolayer(fit1)+
  autolayer(fit2)+
  autolayer(fit3)+
  autolayer(fit4)+
  theme_bw()

Metrics::rmse(sactnetcrc_test, pronsetar1)
Metrics::rmse(sactnetcrc_test, pronsetar2)
Metrics::rmse(sactnetcrc_test, pronsetar3)
Metrics::rmse(sactnetcrc_test, pronsetar4)

autoplot(sactnetcrc_test)+
  autolayer(exp(pronsetar1))+
  autolayer(exp(pronsetar2))+
  autolayer(exp(pronsetar3))+
  autolayer(exp(pronsetar4))+
  theme_bw()
```


### LSTAR

```{r , eval=F}

modlstat1<-lstar(sactnetusd_train, m=1, include = c("const", "trend","none", "both"))     
modlstat2<-lstar(sactnetusd_train, m=2, include = c("const", "trend","none", "both"))  
modlstat3<-lstar(sactnetusd_train, m=3, include = c("const", "trend","none", "both"))  

plot(modlstat1)
plot(modlstat2)
plot(modlstat3)

summary(modlstat1)
summary(modlstat2)
summary(modlstat3)

pronlsetar1<- predict(modlstat1, n.ahead = 7)
pronlsetar2<- predict(modlstat2, n.ahead = 7)
pronlsetar3<- predict(modlstat3, n.ahead = 7)

Metrics::rmse(sactnetcrc_test, pronlsetar1)
Metrics::rmse(sactnetcrc_test, pronlsetar2)
Metrics::rmse(sactnetcrc_test, pronlsetar3)

fit1<-ts(modlstat1$fitted.values,start =c(2011,1),frequency = 12)
fit2<-ts(modlstat2$fitted.values,start =c(2011,1),frequency = 12)
fit3<-ts(modlstat3$fitted.values,start =c(2011,1),frequency = 12)


autoplot(sactnetusd_train)+
  autolayer(fit1)+
  autolayer(fit2)+
  autolayer(fit3)+
  theme_bw()


autoplot(sactnetcrc_test)+
  autolayer(pronlsetar1)+
  autolayer(pronlsetar2)+
  autolayer(pronlsetar3)+
  theme_bw()

```

### Metricas Generales

```{r , eval=F}
Metrics::rmse(sactnetcrc_test,(prontar3))
Metrics::rmse(sactnetcrc_test, (pronsetar2))
Metrics::rmse(sactnetcrc_test, (pronlsetar3))

autoplot(sactnetcrc_test)+
  autolayer((prontar3))+
  autolayer((pronsetar2))+
  autolayer((pronlsetar3))+
  theme_bw()
```
